---
title: "k-means clustering"
author: "Cathy Kiriakos"
date: "October 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## k-means Clusterings
This mini-project is based on the K-Means exercise from ‘R in Action’. Go here for the original blog post and solutions http://www.r-bloggers.com/k-means-clustering-from-r-in-action/.


#Exercise 0: 
install packages
```{r,install.packages}
#install.packages(c("cluster", "rattle.data","NbClust"))
library(rattle.data)
library(NbClust)
library(cluster)
```

Load data and look at first few rows:
```{r, load_data}
data(wine, package="rattle.data")
head(wine)
```

##Exercise 1: 
Remove the first column from the data and scale it using the scale() function:
```{r, remove_cols}
df <- scale(wine[,-1])
```

Now we’d like to cluster the data using K-Means. How do we decide how many clusters to use if you don’t know that already? We’ll try two methods.

## Method 1:
 A plot of the total within-groups sums of squares against the number of clusters in a K-means solution can be helpful. A bend in the graph can suggest the appropriate number of clusters.
 
```{r,wssplot}
wssplot <- function(data, nc=15, seed=1234){
                  wss <- (nrow(data)-1)*sum(apply(data,2,var))
                      for (i in 2:nc){
                set.seed(seed)
                    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
              plot(1:nc, wss, type="b", xlab="Number of Clusters",
                            ylab="Within groups sum of squares")
              wss
       }

wssplot(df)
```
##Exercise 2:
How many clusters does this method suggest? 3
Why does this method work? What’s the intuition behind it? Within group sum of squares dropps off after 3 clusters,, suggesting that a 3-cluster solution may be a good fit to the data.
Look at the code for wssplot() and figure out how it works

##Method 2: 
Use the NbClust library, which runs many experiments and gives a distribution of potential number of clusters.
```{r,nbclust}
library(NbClust)
set.seed(1234)
nc <- NbClust(df, min.nc=2, max.nc=15, method="kmeans")

```
```{r, barplot}
barplot(table(nc$Best.n[1,]),
              xlab="Numer of Clusters", ylab="Number of Criteria",
                    main="Number of Clusters Chosen by 26 Criteria")
table(nc$Best.n[1,])

```
## Exercise 3: 
How many clusters does this method suggest?
14 of 24 criteria provided by the NbClust package suggest a 3-cluster solution.

## Exercise 4: 
Once you’ve picked the number of clusters, run k-means using this number of clusters. Output the result of calling kmeans() into a variable fit.km
```{r,fit}
set.seed(1234)
fit.km <- kmeans(df, centers=3,  nstart=25)
fit.km$size

```

Now we want to evaluate how well this clustering does.
```{r}
table(fit.km$cluster,wine$Type)
```

##Exercise 5: 
using the table() function, show how the clusters in fit.kmclusterscomparestotheactualwinetypesinwineclusterscomparestotheactualwinetypesinwineType. Would you consider this a good clustering?
```{r, table}
table(fit.km$cluster,wine$Type)
```

## Exercise 6:
Visualize these clusters using function clusplot() from the cluster library
Would you consider this a good clustering?
```{r, clusplot}
clusplot(pam(df,3))

```

